---
title: "Data Analysis Introduction Austin Hill"
---

# Installing package

```{r}
install.packages("dslabs")
```

# load dslabs package

```{r}
library("dslabs")
```

# Taking a look at data

```{r}
str(gapminder)
#Summary of data
summary(gapminder)
# determine the type of object gapminder is
class(gapminder)
```

# Install tidyverse package

```{r}
install.packages("tidyverse")
```

# Adding the dplyr that is part of tidyverse

```{r}
library (dplyr)
```

# This code creates a new dataset called africadata that contains all rows from gapminder where the continent is Africa, keeping all columns.

```{r}
africadata <- gapminder %>%
  filter(continent == "Africa")
```

# To test if previous code worked

```{r}
# look at data
str(africadata)
#Summary of data
summary(africadata)
```

# To create two new objects/variables that contain infant_mortality and life_expectancy and one that contains only population and life_expectancy from africa data.

```{r}
# Only infant_mortality and life_expectancy
Africa_infant_life <- africadata %>% select(infant_mortality, life_expectancy)
# Only population and life_expectancy
Africa_pop_life <- africadata %>% select(population, life_expectancy)
```

# To test if previous code worked

```{r}
# look at data
str(Africa_infant_life)
str(Africa_pop_life)
```

```{r}
#Summary of data
summary(Africa_infant_life)
```

```{r}
#Summary of data
summary(Africa_pop_life)
```

# Loading ggplot2 for plotting

```{r}
library (ggplot2)
```

# Infant Mortality vs life Expectancy plot

```{r}
ggplot(Africa_infant_life, aes(x = infant_mortality, y = life_expectancy)) +
  geom_point(color = "blue", size = 2) +
  labs(
    x = "Infant Mortality",
    y = "Life Expectancy",
    title = "Life Expectancy vs Infant Mortality"
  ) +
  theme_minimal()
```

# Life expectancy vs population (log-scaled x-axis) plot

```{r}
ggplot(Africa_pop_life, aes(x = population, y = life_expectancy)) +
  geom_point(color = "darkgreen", size = 2) +
  scale_x_log10() +  # log-scale for x-axis
  labs(
    x = "Population (log scale)",
    y = "Life Expectancy",
    title = "Life Expectancy vs Population"
  ) +
  theme_minimal()
```

# The plot life expectancy vs population that has multiple overlapping sets seems to be from multiple differnt years of data collection.

# Get the unique years with missing infant_mortality data

```{r}
Africainfant_missing_years <- africadata %>%
  filter(is.na(infant_mortality)) %>%  # keep rows with missing infant_mortality
  distinct(year) %>%                    # get unique years
  arrange(year)                         #sort the years

Africainfant_missing_years
```

# Extracting 2000's data from africadata.

```{r}
africa_2000 <- africadata %>%
  filter(year == 2000)
```

# Checking africa_2000

```{r}
str(africa_2000)
summary(africa_2000)
```

# plot for africa 2000 life expectancy vs population

```{r}
ggplot(africa_2000, aes(x = population, y = life_expectancy)) +
  geom_point(color = "darkgreen", size = 2) +
  scale_x_log10() +  # log-scale for x-axis
  labs(
    x = "Population (log scale)",
    y = "Life Expectancy",
    title = "Life Expectancy vs Population"
  ) +
  theme_minimal()
```

# plot for africa 2000 Infant Mortality vs life Expectancy

```{r}
ggplot(africa_2000, aes(x = infant_mortality, y = life_expectancy)) +
  geom_point(color = "blue", size = 2) +
  labs(
    x = "Infant Mortality",
    y = "Life Expectancy",
    title = "Life Expectancy vs Infant Mortality"
  ) +
  theme_minimal()
```

# Running linear regression on life expectancy and infant mortality plot

```{r}
# Fit a linear model: life_expectancy ~ infant_mortality
fit1 <- lm(life_expectancy ~ infant_mortality, data = africa_2000)

# Print a summary of the model
summary(fit1)
```

# This linear regression shows that there is a strong negative relationship between infant mortality and life expectancy among African countries in 2000.Higher infant mortality tend to have lower life expectancy, and this relationship is statistically significant.

# Running linear regression on life expectancy vs population plot

```{r}
# Fit a linear model: life_expectancy ~ population
fit2 <- lm(life_expectancy ~ population, data = africa_2000)

# Print a summary of the model
summary(fit2)
```

# Unlike infant mortality, population size is not a good predictor of life expectancy for African countries in 2000. Any apparent trend is likely just random variation.

# AI was used to help with coding areas I was less familar with.